NOPT048 Notes

### 2019-06-18 misc/study-notes/ndmi011notes.tex
Fixed errors in both notes

\maketitle{NDMI011 Combinatorics and Graph Theory}{Notes by}{Marcel K. Goh {\rm (Prague, Czech Rep.)}}{10 June 2019}

@@

Suppose, towards a contradiction, that $t\in S$. Then there exists a sequence of vertices $v_0,\ldots, v_k\in S$ such that $v_0 = s$, $v_k = t$, and for all $i\in \{0,\ldots, k-1\}$, either
\medskip
\item {a)} $v_iv_{i+1}\in E$ and $c(v_iv_{i+1}) > f(v_iv_{i+1})$, or
\smallskip
\item {b)} $v_{i+1}v_i\in E$ and $f(v_{i+1}v_i) > 0$.
\medskip
For each $i$, define a small real number $\epsilon_i$ given by
$$\epsilon_i = \cases{
    c(v_iv_{i+1}) - f(v_iv_{i+1}) & \hbox{if case (a) holds}\cr
    f(v_{i+1}) & \hbox{if case (b) holds}
}$$
Then we can let $\epsilon = \min\{e_i : 0\leq i\leq k-1\}$. Note that, by construction, $\epsilon > 0$.

@@

So let $M = \{A_1, \ldots, A_t\}$ and for each $A_i\in M$, permute the $n$ columns such that the first row of the resulting Latin square $A'_i$ is $(1,2,3,\ldots,n)$. By our earlier observation, $A'_1,\ldots,A'_k$ are pairwise orthogonal. Now we zoom in on the second row, first column if $A'_i$ for some $i$. Call this cell $k$. First we notice that $k\neq 1$, since the first element of the first row is 1. Then since comparing any two $A'_i$ pairwise, we get all the pairs $(1,1), (2,2), \ldots, (n,n)$ in the first row, each of $2,3,\ldots,n$ can appear in the second row, first column of at most one $A'_i$. So $t = |M|\leq n-1$.\slug

@@

The given bound is tight; there are sequences with length $(r-1)(s-1)$ numbers that neither have an increasing subsequence of length $r$, nor have a decreasing subsequence of length $s$. To construct one, take a grid of $(r-1)(s-1)$ points in the plane and rotate it {\it very} slightly counterclockwise such that no two points are on the same horizontal or vertical line. Then we have a set of points $(x_i, y_i),\ldots(x_n,y_n)$ where $n=(r-1)(s-1)$. Now order the $y$-coordinates {\it by the value of the $x$-coordinates}. Any increasing subsequence can take at most one element from each column, so at most $r-1$, and any decreasing subsequence can take at most one element from each row, so at most $s-1$.

@@

Suppose we have two people who need to communicate over a channel that is ``noisy'' in the sense that errors may be introduced into the message. To detect and solve this issue, checksums may be used. Suppose we have an alphabet $\Sigma = \{0,\ldots,q-1\}$ where $q$ is prime. Let $x$ denote a message $x\in \Sigma^k$ for some positive integer $k$. The {\it checksum} $c$ of $x$ is $\sum_{i=1}^k x_i \pmod{q}$ and we can encode te message as $y\in {\bf F}_q^{k+1}$ where
$$ y_i = \cases{
    x_i & for $i\leq k$ \cr
    c   & for $i = k+1$
}$$
Since $q$ is prime, for any $x,y\in {\bf F}_q^kk$, if $x$ and $y$ differ at {\it exactly one} place, then their checksums will differ. But what happens if more errors occur? Also, is it possible for the receiver to determine where the error occured without asking the sender to resend the message?

@@

A {\it linear} code $C\subseteq {\bf F}_q^n$ is a linear subspace of ${\bf F}_q$, i.e.\  for all $x,y\in C$, $\alpha\in {\bf F}_q$, $\alpha x+y\in C$. There exist $k$ independent vectors $x_1,\ldots, x_k\in {\bf F}_q^n$ such that
$$C = \bigg\{\sum_{i=1}^k \alpha_ix_i : \alpha_i,\ldots \alpha_n\in {\bf F}_q\bigg\}.$$
More succinctly, we can express $C$ using the generator matrix $G\in {\bf F}_q^{n\times k}$. The matrix $G = \matrix{ x_1 & \cdots & x_k }$ and $C = \{G_\alpha : \alpha \in {\bf F}_q^k\}$. (Multiplying any message by $G$ on the left encodes it.)

@@

\proof Define a function $f:\Sigma^n\rightarrow\Sigma^{k-1}$ such that $f(x_1,\ldots, x_n) = (x_1,\dots, x_{k-1})$. As $|\Sigma^{k-1}| = q^{k-1}$ and $|C| = q^k > q^{k-1}$, there are $x,y\in C$, with $x\neq y$ and $f(x) = f(y)$. So $x$ and $y$ can only differ in the last $n-k+1$ entries. Since $d = \Delta(C) \leq \Delta(x,y) \leq n-k+1$, $k\leq n-d+1$.\slug

>>>>

\maketitle{NDMI011 Combinatorics and Graph Theory\footnote{$^*$}{\eightpt Course given by Prof. Andreas Feldmann at Charles University in Prague}}{Notes by}{Marcel K. Goh {\rm (Prague, Czech Rep.)}}{10 June 2019}

@@

Suppose, towards a contradiction, that $t\in S$. Then there exists a sequence of vertices $v_0,\ldots, v_k\in S$ such that $v_0 = s$, $v_k = t$, and for all $i\in \{0,\ldots, k-1\}$, either
\medskip
\item {a)} $v_iv_{i+1}\in E$ and $c(v_iv_{i+1}) > f(v_iv_{i+1})$, or
\smallskip
\item {b)} $v_{i+1}v_i\in E$ and $f(v_{i+1}v_i) > 0$.
\medskip
For each $i$, define a small real number $\epsilon_i$ given by
$$\epsilon_i = \cases{
    c(v_iv_{i+1}) - f(v_iv_{i+1}) & \hbox{if case (a) holds}\cr
    f(v_{i+1}v_i) & \hbox{if case (b) holds}
}$$
Then we can let $\epsilon = \min\{e_i : 0\leq i\leq k-1\}$. Note that, by construction, $\epsilon > 0$.

@@

So let $M = \{A_1, \ldots, A_t\}$ and for each $A_i\in M$, permute the $n$ elements such that the first row of the resulting Latin square $A'_i$ is $(1,2,3,\ldots,n)$. By our earlier observation, $A'_1,\ldots,A'_k$ are pairwise orthogonal. Now we zoom in on the second row, first column if $A'_i$ for some $i$. Call this cell $k$. First we notice that $k\neq 1$, since the first element of the first row is 1. Then since comparing any two $A'_i$ pairwise, we get all the pairs $(1,1), (2,2), \ldots, (n,n)$ in the first row, each of $2,3,\ldots,n$ can appear in the second row, first column of at most one $A'_i$. So $t = |M|\leq n-1$.\slug

@@

The given bound is tight; there are sequences with length $(r-1)(s-1)$ numbers that neither have an increasing subsequence of length $r$, nor have a decreasing subsequence of length $s$. To construct one, take a grid of $(r-1)(s-1)$ points in the plane and rotate it {\it very} slightly counterclockwise such that no two points are on the same horizontal or vertical line. Then we have a set of points $(x_i, y_i),\ldots(x_n,y_n)$ where $n=(r-1)(s-1)$. Now order the $y$-coordinates {\sl by the value of the $x$-coordinates}. Any increasing subsequence can take at most one element from each column, so at most $r-1$, and any decreasing subsequence can take at most one element from each row, so at most $s-1$.

@@

Suppose we have two people who need to communicate over a channel that is ``noisy'' in the sense that errors may be introduced into the message. To detect and solve this issue, checksums may be used. Suppose we have an alphabet $\Sigma = \{0,\ldots,q-1\}$ where $q$ is prime. Let $x$ denote a message $x\in \Sigma^k$ for some positive integer $k$. The {\it checksum} $c$ of $x$ is $\sum_{i=1}^k x_i \pmod{q}$ and we can encode the message as $y\in {\bf F}_q^{k+1}$ where
$$ y_i = \cases{
    x_i & for $i\leq k$ \cr
    c   & for $i = k+1$
}$$
Since $q$ is prime, for any $x,y\in {\bf F}_q^kk$, if $x$ and $y$ differ at {\it exactly one} place, then their checksums will differ. But what happens if more errors occur? Also, is it possible for the receiver to determine where the error occured without asking the sender to resend the message?

@@

A {\it linear} code $C\subseteq {\bf F}_q^n$ is a linear subspace of ${\bf F}_q$, i.e.\  for all $x,y\in C$, $\alpha\in {\bf F}_q$, $\alpha x+y\in C$. There exist $k$ independent vectors $x_1,\ldots, x_k\in {\bf F}_q^n$ such that
$$C = \bigg\{\sum_{i=1}^k \alpha_ix_i : \alpha_i,\ldots \alpha_n\in {\bf F}_q\bigg\}.$$
More succinctly, we can express $C$ using the generator matrix $G = \left[\matrix{ x_1 & \cdots & x_k }\right]\in {\bf F}_q^{n\times k}$. Multiplying any message by $G$ on the left encodes it and $C = \{G_\alpha : \alpha \in {\bf F}_q^k\}$.

@@

\proof Define a function $f:\Sigma^n\rightarrow\Sigma^{k-1}$ such that $f(x_1,\ldots, x_n) = (x_1,\dots, x_{k-1})$. As $|\Sigma^{k-1}| = q^{k-1}$ and $|C| = q^k > q^{k-1}$, there are $x,y\in C$, with $x\neq y$ and $f(x) = f(y)$. So $x$ and $y$ can only differ in the last $n-k+1$ entries. Since $d = \Delta(C) \leq \Delta(x,y) \leq n-k+1$, we get $k\leq n-d+1$.\slug
